---
title: "MCA"
author: "Philipp Bayer"
date: "2022-09-14"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

MCA is an extension of CA that works better when the dataset is larger. Perhaps most famously used in the social sciences in *Bourdieu, P. "La distinction." Revue Philosophique de la France Et de l 172.1 (1982).*

We already did some EDA in the [CA part](CA.html) so we will not repeat that here, rather we'll jump straight into the MCA. We will use the same packages as CA - FactoMineR and factoextra. See tutorials like https://rpubs.com/gaston/MCA  or https://cran.r-project.org/web/packages/factoextra/readme/README.html or http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/114-mca-multiple-correspondence-analysis-in-r-essentials/ 

# Setup 

```{r}
library(FactoMineR)
library(factoextra)
library(tidyverse)
```

```{r}
df <- data.frame(readxl::read_xlsx('./data/MCA_Macropods_Data.xlsx', sheet = 'Macropods_Complete ind'))
head(df)
```

# MCA

```{r}
res.mca <- MCA(df, graph = FALSE, ncp = 5)
summary(res.mca)

eig.val <- get_eigenvalue(res.mca)
eig.val
```

Now we have suddenly WAY more rows - 223 - so our cutoff differs. 1 / 222 * 100 = 0.45
```{r}
fviz_screeplot(res.mca,addlabels=T) + 
  geom_hline(yintercept=0.45,linetype=2,color="red")
```

As you can see, MCA has a MUCH lower explained variances than CA, but our cutoff changed?

## Plotting

```{r}
fviz_mca_biplot(res.mca, 
               repel = TRUE, # Avoid text overlapping (slow if many point)
               ggtheme = theme_minimal())
```

Oh. This particular dataset is much larger than the one I used for CA - investigating.

## Contributions

```{r}
fviz_contrib(res.mca, choice="var",axes=1:2, top = 20)
```

```{r}
fviz_mca_var(res.mca, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE, # avoid text overlapping (slow)
             ggtheme = theme_minimal()
             )

```


And by individual?

```{r}
fviz_mca_ind(res.mca, col.ind = "cos2", 
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE, # Avoid text overlapping (slow if many points)
             ggtheme = theme_minimal())
```

There are some weird outlier individuals. Hm.
